---
title: "Hands-on Exercise 1"
author: "SHIH, HUA-HSUAN"
---

# Getting Started

## Install and launching R packages

The code chunk below uses p_load() of pacman package to check if tidyverse packages are installed in the computer. If they are, then they will be launched into R.

```{r}
pacman::p_load(tidyverse)
```

## Importing the data

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

# Plotting a simple bar chart

```{r}
ggplot(data = exam_data,
       aes(x = RACE)) +
  geom_bar()
```

# Histogram with R graphic

```{r}
hist(exam_data$MATHS)
```

# Histogram with ggplot

```{r}
ggplot(data=exam_data, aes(x = MATHS)) +
  geom_histogram(bins=10, 
                 boundary = 100,
                 color="black", 
                 fill="grey") +
  ggtitle("Distribution of Maths scores")
```

# 1.7 Essential Grammatical Elements in ggplot2: geom

## 1.7.2 Geometric Objects: geom_dotplot點點圖

```{r}
ggplot(data=exam_data, aes(x=MATHS))+
  geom_dotplot(binwidth = 2.5,
               dotsize = 0.5) + 
  scale_y_continuous(NULL,
                     breaks = NULL)
```

## 1.7.3 Geometric Objects: geom_histogram()直方圖

```{r}
ggplot(data=exam_data,
       aes(x = MATHS)) +
  geom_histogram()
```

## 1.7.5 Modifying a geometric object by changing aes() 摺疊直方圖

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           fill = GENDER)) +
  geom_histogram(bins=20, 
                 color="grey30")
```

## 1.7.6 Geometric Objects: geom-density() 折線圖

```{r}
ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_density()           
```

```{r}
ggplot(data=exam_data, 
       aes(x = MATHS, 
           colour = GENDER)) +
  geom_density()
```

## 1.7.7 Geometric Objects: geom_boxplot箱型圖

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS,       
           x= GENDER)) +    
  geom_boxplot()            
```

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, 
           x= GENDER)) +
  geom_boxplot(notch=TRUE)
```

## 1.7.8 Geometric Objects: geom_violin 小提琴圖

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, 
           x= GENDER)) +
  geom_violin()
```

## 1.7.9 Geometric Objects: geom_point()散點圖

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point()   
```

## 1.7.10 geom objects can be combined 組合圖

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, 
           x= GENDER)) +
  geom_boxplot() +                    
  geom_point(position="jitter", 
             size = 0.5)        
```

# 1.8 Essential Grammatical Elements in ggplot2: stat

## 1.8.2 Working with stat - the stat_summary() method 加上平均值

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, x= GENDER)) +
  geom_boxplot() +
  stat_summary(geom = "point",       
               fun = "mean",         
               colour ="red",
               size=4)
```

## 1.8.3 Working with stat - the geom() method 加上平均值

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, x= GENDER)) +
  geom_boxplot() +
  geom_point(stat="summary",        
             fun="mean",           
             colour="red",          
             size=4)          
```

## 1.8.4 Adding a best fit curve on a scatterplot 趨勢線

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, y=ENGLISH)) +
  geom_point() +
  geom_smooth(size=0.5)
```

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              linewidth=0.5)
```

# 1.9 Essential Grammatical Elements in ggplot2: Facets

## 1.9.1 Working with facet_wrap()多張圖

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS)) +
  geom_histogram(bins=20) +
    facet_wrap(~ CLASS)
```

## 1.9.2 facet_grid() function

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS)) +
  geom_histogram(bins=20) +
    facet_grid(~ CLASS)
```

# 1.10 Essential Grammatical Elements in ggplot2: Coordinates

## 1.10.1 Working with Coordinate 圖表轉向

```{r}
ggplot(data=exam_data, 
       aes(x=RACE)) +
  geom_bar() +
  coord_flip()
```

## 1.10.2 Changing the y- and x-axis range 交換x,y軸

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, size=0.5)
```

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100))
```

# 1.11 Essential Grammatical Elements in ggplot2: themes

## 1.11.1 Working with theme 修改底色

```{r}
ggplot(data=exam_data, 
       aes(x=RACE)) +
  geom_bar() +
  coord_flip() +
  theme_gray()
```

```{r}
ggplot(data=exam_data, 
       aes(x=RACE)) +
  geom_bar() +
  coord_flip() +
  theme_classic()
```

```{r}
ggplot(data=exam_data, 
       aes(x=RACE)) +
  geom_bar() +
  coord_flip() +
  theme_minimal()
```
